---
title: "linear regression"
auhtor: "Timothy Morris, PhD" 
output: html_document
---

# this tutorial will run through the basics of simple (1 independant variable and 1 continuous dependant variable) and mulitple (several independant varibales and one continuous dependant varibale). We will load in a dataset, 'tidy' it, run basic descriptive statistics, check for the assumptions of linear regression, remove outliers, interpret the output and plot the results. 

# load libraries 
```{r setup, include=FALSE}
library(tidyverse)
```
# import the data
```{r}
setwd('/Users/timothymorris/Library/Mobile Documents/com~apple~CloudDocs/Studies/statistics') # this will change depending on  where you pulled the github repo to on your local computer. 

d<-read.csv('FK_example.csv',sep=',',header = T)

library(tidyverse) 
d <- as_tibble(d) %>%
    rename(ID = PartID,
          age = Pre_Age,
          sex = Pre_Sex,
          ses = Pre_SES,
          pr.vo2.pc = Pre_VO2max_perc,
          po.vo2.pc = Post_VO2max_perc,
          pr.congrt = Pre_FL_CompCong_MeanRT,
          po.congrt = Post_FL_CompCong_MeanRT,
          pr.inconrt = Pre_FL_CompIncon_MeanRT,
          po.inconrt = Post_FL_CompIncon_MeanRT) %>%
    #filter(filter ==1)  %>% #  Filter (e.g., based on filter variable created in SPSS) 
    mutate(study = factor(FK1_FK2)) %>%
    mutate(wave = factor(Wave)) %>% 
    #mutate (wtclass = recode_factor(wtclass, '1' = "Normal Weight", '2' = "Overweight", '3' = "Class 1 & 2 Obese")) %>%
    mutate(ses = factor(ses)) %>%
    mutate(ID = factor(ID)) %>%
    mutate(sex = recode_factor(sex, '0' = "female", "1" = "male"))
```

