---
title: "Linear regression workshop"
author: "Timothy P Morris"
date: "9/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

1. Import data 
2. Tidy data 
3. Summarize and visualize your data
4. Check for assumptions (visually and statistically) 
5. Transform your data if necessary (outlier removal, transformations)
6. Run your model
7. Plot your model 
8. Extract standardized coefficients 
9. Confounders 
10. Interactions and simple slopes (continuous continuous and continuous categorical interactions)
11. Bootstrapping
12. 

#load libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(psych)
library(corrr)
library(nlme)
library(ggstatsplot)
library(glmnet)
library(fpc)
```
# import the data
```{r}
#load and sort data data
setwd("/Users/timothy/iCloud/1.Tims_Team_Lab/statistics") # set working dir

load(file = "middle.Rdata")

# test this change 
```
# summarize and visualize your data
```{r}
#summarise the whole df
#library(psych)
describe(middle)
#create new df and do corr plot and histograms
demos <- middle %>%
  select(bmi,
         sex,
         educ,
         age,
         SES,
         pds)
print(kable(middleniih))
pairs.panels(demos)
#again for smaller n of vars
nih <- middle %>%
  select(walk_enduro, 
         gait_speed,
         dexterity,
         grip_strength)
pairs.panels(nih)
#plot corr plot with fun network plot 
#library(corrr)
middle %>%
  select(age,bmi:grip_strength) %>%
  na.omit() %>%
  correlate() %>%
  rearrange() %>%
  network_plot(colours = c("orange", "white", "midnightblue"))
#summarise by group or groups of vars
# Group by a single variable (in this case by sex)
middle %>%
  group_by(sex) %>%
  summarise_all(mean)

# group by several variables (group by sex and physical activity category and summarise bmi)
middle %>%
group_by(sex, PA_cat12) %>%
  summarise(
    n = n(),
    mean = mean(bmi),
    sd = sd(bmi)
  )

# take a closer look at some of the distributions with density plots

# Histogram overlaid with kernel density curve
ggplot(middle, aes(x=bmi)) + 
    geom_histogram(aes(y=..density..),      # Histogram with density instead of count on y-axis
                   binwidth=.5,
                   colour="black", fill="white") +
    geom_density(alpha=.2, fill="#FF6666")  # Overlay with transparent density plot

# overlaid hist by groups
# Overlaid histograms
middle$sex <- as.factor(middle$sex) # convert sex to cat
ggplot(middle, aes(x=bmi, fill=sex)) +
    geom_histogram(binwidth=.5, alpha=.5, position="identity")

```
# run simple and multiple linear regressions
```{r}
#model
lmodel <- summary(lm(cog_fluid ~ bmi, data = middle))
lmodel
```
# plot your model
```{r}
# note- run chunk x (last chunk first to create function)
scatterplot(middle, bmi, cog_fluid, "bmi", "gait speed") # x is predictor (independant) and y is dependent. 
```
# check linear regression assumptions: (1. indepedence 2. Linearity 3. Homogeneity of variance 4. normality of residuals)
# function "diagnostics" needs to be provided with x and y as 'd$x, d$y' 
```{r}

# see this website for a good explanation: https://data.library.virginia.edu/diagnostic-plots/ 

#visually check for asusmptions using this function 
diagnostics <-function(y,x) {
par(mfrow=c(2,2)) #plots 4 plots to visually check assumptions 
plot(lm({{y}} ~ {{x}}))
}

diagnostics(middle$bmi, middle$cog_fluid)

#1 Linearity is only really a problem in time series data- but gets to basic algebra- is the relatiionship between two variables best described as a straight line ð‘¦=ð‘Ž+ð‘ð‘¥ 

#2 Independence is more if a theoretical issue (are these two variables independent of each other)

#3 Homoscedacticity Briefly summarizing the information from the websites above, heteroscedasticity does not introduce a bias in the estimates of your coefficients. However, given heteroscedasticity, you are not able to properly estimate the variance-covariance matrix. Hence, the standard errors of the coefficients are wrong. This means that one cannot compute any t-statistics and p-values and consequently hypothesis testing is not possible. 

#4 Normality of the residuals- similar to heteroscedasticity, if this assumption is violated then the error across the model (predictivness of x on y) is not constant across different levels of x- i.e., the predictor means different things are different levels of the dependent variable- this assumption is not such a big issue with large datasets. 
# In this example we can see that the normality of the residuals is violated. Given the large sample size its probably ok, but lets transform the dep[endant variable anyway to see what it does.

# If we violated the assumption of homoscedasticity, we could try and transform our response/dependent variable. 

# OUTLIERS - the residual vsa leverage plot shows us influential outliers using Cook's distance of 0.5 and tells us if any data points are significantly influential over the models coefficients. 

# formally check - you can formally (instead of visually) check for two of these assumptions using a Shapiro-Wilk Test (normality of residuals) and 

# transform predictor using log10

middle$bmilog <- log10(middle$bmi)
hist(middle$bmilog)

logfmodel <- summary(lm(cog_fluid ~ bmilog, data = middle))
logfmodel
diagnostics(middle$bmilog, middle$cog_fluid)

```
# run multiple linear regresison checking for multicolinearity
```{r}

lmmodel <- lm(cog_fluid ~ dexterity + bmilog, data = middle)
summary(lmmodel)
install.packages("car")
library(car)
vif(lm(cog_fluid ~ dexterity + bmilog + sex, data = middle))# if larger than 10 
 
vif(lm(bmi ~ cog_fluid + cog_total + cog_crystal + sex, data = middle))# if larger than 10 

```
# extract standardized beta coefficients 
```{r}
install.packages("QuantPsyc")
library(QuantPsyc)
lm.beta(lmmodel)
```
# control for multiple comparison corrections
```{r}
# note, this is for when we run several independent models - for corrections of ANOVA or the like Bonferoni, Tukey HSD are preffered. 

# create a df with two columns, one for the number of comparisons and the other for the p value - and rank them.

pvals <- data.frame(number = c(1,2,3,4,5),
                 pvalue = c(0.001, 0.02, 0.05, 0.65, 0.77)
                 )

# run fdr

pvals$fdrval<-p.adjust(pvals$pvalue, method="fdr")
pvals
```

# run multilevel model accounting for site 
```{r}
summary(lme(cog_fluid~ dexterity + age + sex + SES + pds, random=~1|site, data=middle)) # 2 clusters

```
# Plotting 
```{r}

# for an amazing resource of R plots see: http://r-graph-gallery.com/ 

# Run lines 2-18 once to create function, then call function using scatterplot(dataframe, x, y, "ylab", "xlab")
library(rlang)
library(ggplot2)
scatterplot <- function(dataframe, x, y, xlab, ylab) {
  ggplot(data = dataframe, aes(x={{x}}, y = {{y}})) + 
    geom_point(color =  "#663399") + 
    theme(
      # Hide panel borders and remove grid lines
      panel.border = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      # Change axis line
      axis.line = element_line(colour = "black")
    ) +
    labs(y=ylab, x=xlab) +
    stat_smooth(method = "lm", col = "#cc8400")
}
```


